---
layout: tutorial
title: Test covariates
title_id: test_covariates
---

### LRT

`SOLAR` performs likelihood ratio test (LRT) 
to evaluate the significance of covariates in a model.
`covtest` argument says to conduct the LRT tests for all covariates.
This option is disabled by default (to save calculation time).

```{r}
mod1 <- solarPolygenic(trait1 ~ age + sex, dat30)
mod1$cf
```

For the next model the test results are collected into `cf` slot
of the model object.

```{r}
mod2 <- solarPolygenic(trait1 ~ age + sex, dat30, covtest = TRUE)
mod2$cf
```

The `SOLAR` command was the following.

```{r}
mod2$solar$cmd
```

In particular, `-screen -all` options for `polygenic` command do the job.
These two options say to screen the covariates, although all covariates will be saved (not screened).
That leads to the tests without covariate screening.

### Model files

`SOLAR` generates files to store the modeling results.

```{r}
names(mod2$solar$files$model)
```

Files `noage.mod` and `nosex.mod` are two polygenic models without one of the two covariates,
which in turn were compared to the null model with all covariates to get the LRT statistics.
