---
title: "SOLAR command fix"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---


## Include

```{r inc}
library(solarius)
```

## Parameters

```{r par}
dir1 <- "solar1"
dir2 <- "solar2"
dir3 <- "solar3"
```

```{r dir_create}
dir.create(dir1, showWarnings = FALSE)
dir.create(dir2, showWarnings = FALSE)
dir.create(dir3, showWarnings = FALSE)
```


## Example 1

```{r ex1}
dat <- loadMulticPhen()
df2solar(dat, dir1)

cmd1 <- c("load pedigree dat.ped", 
  "load phenotypes dat.phe",
  "trait trait1",
  "covariate",
  "polygenic",
  "quit")

ret1 <- solar(cmd1, dir1)
grep("H2r", ret1, value = TRUE)

cmd2 <- c("load model trait1/null0",
  "covariate age",
  "polygenic",
  "quit")
ret2 <- solar(cmd2, dir1)
grep("H2r", ret2, value = TRUE)
```

## Example 2

Compute two models:

* `trait1/null0`: no covariates, free `h2r`
* `constrained.mod`: `age` covariate, constrained `h2r` to that value in the previous model

```{r ex2}
dat <- loadMulticPhen()
df2solar(dat, dir2)

cmd1 <- c("load pedigree dat.ped", 
  "load phenotypes dat.phe",
  "trait trait1",
  "covariate",
  "polygenic",
  "quit")

ret1 <- solar(cmd1, dir2)
grep("H2r", ret1, value = TRUE)

cmd2 <- c("load model trait1/null0",
  "covariate age",
  "fix h2r",
  "maximize", "save model trait1/constrained",
  "quit")
ret2 <- solar(cmd2, dir2)
grep("H2r", ret2, value = TRUE)
```

Compare likelihoods:

```{r read_constrained}
tail(readLines("solar2/trait1/null0.mod"), 1)
tail(readLines("solar2/trait1/constrained.mod"), 1)
```

Make sure that the first likelihood is the same as that computed by `solarPolygenic` function:

```{r mod1}
mod1 <- solarPolygenic(trait1 ~ 1, dat)
mod1$lf
```

Compare h2r:

```{r compare_h2r}
grep("h2r", readLines("solar2/trait1/null0.mod"), value = TRUE)[1]
grep("h2r", readLines("solar2/trait1/constrained.mod"), value = TRUE)[1]
```

## Example 3

Compute two models:

* `trait1.trait2/null0`: no covariates, free variance components
* `constrained.mod`: `age` covariate, constrained free variance components

```{r ex3}
dat <- loadMulticPhen()

data(dat30)
dat <- dat30

df2solar(dat, dir3)

cmd0 <- c("load pedigree dat.ped", "load phenotypes dat.phe",
  "trait trait1 trait2", "covariate age",
  "polygenic", 
  "quit")

t0 <- system.time(ret0 <- solar(cmd0, dir3))

cmd1 <- c("load model trait1.trait2/null0",
  "covariate sex sex",
  "polymod", "maximize", "save model trait1.trait2/constrained1",
  "quit")

t1 <- system.time(ret1 <- solar(cmd1, dir3))

cmd2 <- c("load model trait1.trait2/null0",
  "covariate age sex",
  "constraint delete_all",
  "fix h2r(trait1)", "fix h2r(trait2)",
  "fix e2(trait1)", "fix e2(trait2)",  
  "fix rhoe", "fix rhog",
  "maximize", "save model trait1.trait2/constrained2",
  "quit")

t2 <- system.time(ret2 <- solar(cmd2, dir3))

cmd3 <- c("load model trait1.trait2/null0",
  "covariate age sex",
  "constraint delete_all",
  "fix h2r(trait1)", "fix h2r(trait2)",
  "fix e2(trait1)", "fix e2(trait2)",  
  "fix rhoe", "fix rhog",
  "fix mean(trait1)", "fix mean(trait2)",  
  "fix sd(trait1)", "fix sd(trait2)",  
  "fix bage(trait1)", "fix bage(trait2)",  
  "maximize", "save model trait1.trait2/constrained3",
  "quit")

t3 <- system.time(ret3 <- solar(cmd3, dir3))

cmd4 <- c("load model trait1.trait2/null0",
  "covariate age sex",
  "constraint delete_all",
  "fix h2r(trait1)", "fix h2r(trait2)",
  "fix e2(trait1)", "fix e2(trait2)",  
  "fix rhoe", "fix rhog",
  "fix mean(trait1)", "fix mean(trait2)",  
  "fix sd(trait1)", "fix sd(trait2)",  
  "fix bage(trait1)", "fix bage(trait2)",  
  "option MaxIter 0", "option StandErr 0",
  #"option MaxStep 0", "option MaxCliffs 0",
  "maximize", "save model trait1.trait2/constrained4",
  "quit")

t4 <- system.time(ret4 <- solar(cmd4, dir3))
```

```{r}
c(t0[3], t1[3], t2[3], t3[3], t4[3])
```

```{r}
round(c(t0[3], t1[3], t2[3], t3[3], t4[3]) / t3[3], 2)
```


```{r}
round(c(t0[3], t1[3], t2[3], t3[3], t4[3]) / t4[3], 2)
```
